### 2019-01-17 BSS7782.tex
updated tex file

%%
%% Beginning of file 'sample62.tex'
%%
%% Modified 2018 January
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.2 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass{aastex62}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex62}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Preferred style for Research Notes which are by design 
%%                lacking an abstract and brief. DO NOT use \begin{abstract}
%%                and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

\newcommand{\MSun}{\mbox{M$_\odot$}}
\newcommand{\RSun}{\mbox{R$_\odot$}}
\newcommand{\LSun}{\mbox{L$_\odot$}}

\def\apgt{\ {\raise-.5ex\hbox{$\buildrel>\over\sim$}}\ }
\def\aplt{\ {\raise-.5ex\hbox{$\buildrel<\over\sim$}}\ }
%%\def\figure{\ {figure\,}}

\def\Simon#1{{\bf {\color{red}[#1 -- Simon]}}}
\def\simon#1{{\bf {\color{red}[#1 -- Simon]}}}
\def\del#1{{\bf {\sout{#1}}}}
\def\replace#1#2{{\bf {\sout{#1} $\rightarrow$ {\bf #2}}}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{January 1, 2018}
\revised{January 7, 2018}
\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{Compact Binaries of BS Twins from Stellar Triples}
\shortauthors{Portegies Zwart \& Leigh}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{A Triple Origin for Twin Blue Stragglers in Close Binaries}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.2 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.2 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Nathan W. C. Leigh}
\email{nleigh@amnh.org}

\author{Simon Portegies Zwart}
\affiliation{Leiden Observatory \\
Leiden University \\
PO Box 9513, 2300 RA \\
Leiden, the Netherlands}

\author{Nathan W. C. Leigh}
\affil{American Museum of Natural History \\
Department of Astrophysics \\
79th Street at Central Park West \\
New York, NY 10024-5192, USA}
\affil{Stony Brook University \\
Department of Physics and Astronomy\\
Stony Brook, NY 11794-3800, USA}
\affil{Departamento de Astronom\'ia \\ 
Facultad de Ciencias F\'isicas y Matem\'aticas \\ 
Universidad de Concepci\'on \\ 
Concepci\'on, Chile}

%\author{Simon Portegies Zwart}
%\affiliation{Leiden Observatory \\
%Leiden University \\
%PO Box 9513, 2300 RA \\
%Leiden, the Netherlands}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%\newcommand{\mnras}{\textrm{MNRAS}}
%\newcommand{\apj}{\textrm{ApJ}}
%\newcommand{\aap}{\textrm{A\&A}}
%\newcommand{\apjl}{\textrm{ApJ}}

\begin{abstract}

We propose a formation mechanism for twin blue stragglers (BSs) in compact
binaries that involves mass transfer from an evolved outer tertiary
companion on to the inner binary via a circumbinary disk.  We apply
this scenario to the observed double BS system Binary 7782
in the old open cluster NGC 188, and show that its observed properties
are naturally reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin BSs: (1) For the outer tertiary orbit, the initial orbital
period should lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$
1100 days, assuming initial masses for the inner binary components of
$m_{\rm 1} = 1.1$ M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an
outer tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  After
Roche-lobe overflow, the outer star turns into a white dwarf (WD) of
0.43 to 0.54\,\MSun. There is a correlation between the mass of this
WD and the outer orbital period: more massive WDs will be on wider orbits.  (3) The rotational axes of both 
BSs will be aligned with each other and the
orbital plane of the outer tertiary WD. (4) The BSs 
will have roughly equal masses, independent of their initial
masses (since the initially lower mass star accretes the most).  The dominant accretor should, therefore, be enriched
more effectively by the accreted material.  As a result, one of the
BSs will appear to be more enriched by either He, C and O or by s-process elements, depending on 
if the donor started to overflow its Roche lobe on, respectively, the red
giant or asymptotic giant branch.  (5) Relative to old dense clusters with high-velocity dispersions, twin BSs in close binaries formed from the
proposed mechanism should be more frequent in the Galactic field and
younger open clusters with ages $\lesssim$ 4-6 Gyr, since then the
donor will have a radiative envelope. (6) the orbit of the binary BS
will have a small semi-major axis (typically $\aplt 0.3$\,au) and be close to circular ($e
\aplt 0.2$).

\end{abstract}

\keywords{stars: blue stragglers -- binaries: general -- globular clusters: general -- scattering}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{intro}

Blue straggler stars are brighter and bluer than the
main-sequence (MS) turn-off in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,2014ApJ...782...49S}.  Two primary
channels for BS formation have been proposed: mass transfer from an
evolved donor on to a MS star in a binary star system
\citep[e.g.][]{1964MNRAS.128..147M,1997A&A...328..143P,2009Natur.457..288K,2011MNRAS.410.2370L,2011Natur.478..356G},
and direct stellar collisions involving MS stars possibly
mediated via binaries
\citep[e.g.][]{1975AJ.....80..809H,1997A&A...328..130P,2007ApJ...661..210L,2013MNRAS.428..897L,2013MNRAS.429.1221H,2018arXiv181100058P}.
Other possible, albeit related, formation mechanisms include mergers
of close MS-MS binaries \cite{2018arXiv181100058P},
and mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{2009ApJ...697.1048P}.

In spite of these specific predictions for the expected properties of
BSs formed from each of the above production mechanisms, many BSs
exist with observed properties that defy these simple scenarios.  For
example, in the old open cluster (OC) M67, there lurks a candidate triple
system that is posited to host two BSs
\citep{2001A&A...375..375V,2003AJ....125..810S}.  The observations
suggest that the outer tertiary is itself a BS, with a mass $\sim$ 1.7
M$_{\odot}$ and orbiting the inner binary with a period of $\sim
1188.5$ days \citep{2003AJ....125..810S}.  The inner binary has a
period of only $\sim 1.068$ days \citep{2001A&A...375..375V}, and
hosts a BS of mass $\sim 2.52$ M$_{\odot}$.  In order to reproduce the total system mass 
we require at least five stars \citep{2011MNRAS.410.2370L}.  This is strongly indicative of a
dynamical origin for the system, and a single direct interaction
involving a binary and a triple that resulted in two separate
collisions is the most probable explanation for its origin
(instead of back-to-back direct binary-binary interactions)
\citep{2004MNRAS.350..615G,2011MNRAS.410.2370L}.

Even more curious, there exists in the old OC NGC 188 a
double BS binary, called Binary 7782.  The BS
population in NGC 188 has a bi-modal period-eccentricity distribution.
As discussed in \citet{2011MNRAS.410.2370L}, this could be hinting at
a triple origin for at least some subset of the total BS population.
As for Binary 7782, \citet{2009Natur.462.1032M} observed a compact and
mildly eccentric (i.e., $e \sim 0.1$) binary star system with an
orbital period of $\sim$ 10 days hosting two roughly equal-mass BSs.  During a given binary-binary interaction, the probability
that not one but two direct (MS-MS) collisions will occur is less than
$10^{-2}$
\citep{1989AJ.....98..217L,2011MNRAS.410.2370L,2012MNRAS.425.2369L}.
Plus, binaries with collision products typically have relatively long
orbital periods \cite{2011Sci...334.1380F}. Dynamically, it is
difficult to form a short-period binary composed of two collision
products during a collisional interaction in a star cluster
\citep{2011Sci...334.1380F}.  So, how did Binary 7782 form?

We propose a formation channel for Binary 7782, and compact double BS
binaries in general, which involves mass transfer from an outer
tertiary companion on to an inner binary composed of two MS 
stars.  In section~\ref{sect:dyn}, we constrain the range of initial
(i.e., pre-mass transfer) orbital parameters for a hypothetical outer
tertiary companion, using a combination of dynamical and stellar
evolution-based constraints.  In Section~\ref{sims} we present the
numerical simulations used to study the mass transfer process in this triple
system. We adopt orbital parameters that, according to our expectations,
are most promising for the progenitors of the twin BS 7782.  The calculations are
performed using the Astrophysical Multipurpose Software Environment
(\texttt{AMUSE} for short) \cite{AMUSE} with a combination of stellar
evolution, hydrodynamical and gravitational simulations.  with these
calculations we further constrain the possible range of initial
parameters that naturally lead to twin BSs with orbital parameters
similar to the 7782 system.  We summarize and discuss the implications
of our results for compact double BS binaries and, more generally,
mass transfer in stellar triples in Section~\ref{sect:discussion}.

\section{Constraints on the present-day orbital parameters for a hypothesized
         tertiary companion in the compact BS Binary 7782} \label{sect:dyn}

In our scenario, we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited by a tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  For clarity we 
assume both orbits, the inner as well as the outer, to have negligible
eccentricity and low inclination.  These assumptions are also
supported by the population of observed low-mass triples
\cite{2010yCat..73890925T,moe18}.  This initial configuration for our
assumed formation scenario for Binary 7782, described below, is
depicted in figure~\ref{fig:fig1}.

\begin{figure}[ht!]
\plotone{fig1.eps}
\caption{Cartoon depiction of our proposed scenario for the formation
  of Binary 7782, specifically mass transfer from an evolved outer
  tertiary companion on to a compact inner binary via a circumbinary
  disk.  The outer tertiary component has mass m$_{\rm 3}$, whereas
  the inner binary components have masses m$_{\rm 1}$ and m$_{\rm 2}$.
  The inner and outer orbital separations are denoted by,
  respectively, a$_{\rm in}$ and a$_{\rm out}$.  The circularization
  radius of the accretion stream is denoted a$_{\rm c}$, as calculated
  via Equation~\ref{eqn:ac}, and marks the mean separation of the
  circumbinary disk.
\label{fig:fig1}}
\end{figure}

According to our scenario $m_3 > m_1 > m_2$ and the outer orbit is
sufficiently small that the tertiary star is filling its Roche lobe
and transfers mass to the inner binary before it leaves the asymptotic
giant branch. We constrain the inner orbit by requiring the triple
system to be dynamically stable, for which we adopt eq.\,1 in
\cite{1999ASIC..522..385M,}.  While transferring mass, the accretion
stream gathers around the inner binary at the circularization radius
a$_{\rm c}$, and forms a circumbinary disk
\citep{2002apa..book.....F}.  Using conservation of angular momentum,
we equate the specific angular momentum of the accreted mass at the
inner Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the centre of mass corresponding to the tertiary defined by the
Roche lobe is given by eq.\,2 in \citep{1983ApJ...268..368E}.
Combining eq.\,2 in \citet{1983ApJ...268..368E} (with mass ratio q
$=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
eq.~\ref{eqn:specangmom1}, we solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that a$_{\rm in} <$ a$_{\rm c}$.

Figure~\ref{fig:fig2} shows the parameter space in the P$_{\rm
  out}$-P$_{\rm in}$-plane for Binary 7782.  We assume initial
component masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and $m_{\rm 2} =
0.9$ M$_{\rm \odot}$ for the inner binary components, and $m_{\rm 3} =
1.4 $M$_{\rm \odot}$ for the outer tertiary.  We compare the
circularization radius to the semi-major axis of the inner binary, for
which we require $a_{\rm c} > a_{\rm in}$, after folding in all
constraints from the requirements for dynamical stability (listed in
the caption of figure~\ref{fig:fig2}), and the assumption of an outer tertiary that is Roche
lobe-filling.  
%\simon{Why do you consider it important to mention that the orbit is hard, and for what environment did you consider this? - NL: Deleted.}  
Note that the range of
plotted orbital periods P$_{\rm in}$ corresponding to a contact state
for the inner binary lies outside the range of plotted values for
$P_{\rm in}$ (for components with radii of 1 R$_{\odot}$),
since it does not contribute to constraining the outer orbital
properties.  The thick horizontal solid red line shows the allowed
range of outer semi-major axes, after folding in all of 
the aforementioned criteria.  These constraints result in a rather narrow
range of initial conditions for the outer orbit, namely 2.2 $\times$
10$^{2}$ days $\le$ P$_{\rm out}$ $\le$ 1.1 $\times$ 10$^3$ days.
%, for
%which the tertiary star overfills its Roche lobe, the triple is (at
%least initially) dynamically stable and the accretion stream forms a
%gaseous ring around the inner binary.

\begin{figure}[ht!]
\plotone{fig2.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782
  before Roche-lobe overflow.  The solid diagonal black line shows the
  period corresponding to the circularization radius a$_{\rm c}$ for
  the mass transfer stream coming from the outer star (i.e., at the
  onset of mass transfer), assuming a mass ratio of q $=$ 0.7 and
  component masses of m$_{\rm 1} =$ 0.9 M$_{\odot}$, m$_{\rm 2} =$ 1.1
  M$_{\odot}$ and m$_{\rm 3} =$ 1.4 M$_{\odot}$.  We assume initial
  component masses of m$_{\rm 1}$ = 1.1 M$_{\rm \odot}$ and m$_{\rm
    2}$ = 0.9 M$_{\rm \odot}$ for the inner binary components, and
  m$_{\rm 3}$ is computed for the outer tertiary according to our
  assumed mass ratio (with our fiducial case corresponding to q $=$
  0.7).  We assume completely conservative mass transfer for this
  exercise, and a final mass for the outer tertiary of 0.6 M$_{\rm
    \odot}$ once it has become a WD.  The dashed diagonal black line
  shows a rough criterion for dynamical stability in the triple,
  approximately following \citet{mardling99} (i.e., a$_{\rm in} <$
  0.1a$_{\rm out}$ is required for long-term dynamical stability in
  equal-mass co-planar triples).  The vertical solid red lines show
  the outer orbital periods corresponding to the hard-soft boundary
  assuming central velocity dispersions of $\sigma =$ 1, 5 and 10 km
  s$^{-1}$.  The vertical dashed black lines show the maximum outer
  orbital period P$_{\rm out}$ for which the outer tertiary companion
  is Roche lobe-filling, assuming a stellar radius of R$_{\rm 3} =$
  200 R$_{\odot}$ (which corresponds to the maximum stellar radius
  reached on the AGB for the range of tertiary masses of interest to
  us; see figure~\ref{fig:tertiarymass_vs_size}).  The horizontal dashed red line shows the observed orbital
  period for Binary 7782, using its observed orbital period and our
  assumed final inner companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2}
  =$ 1.4 M$_{\odot}$).  Finally, the thick solid horizontal red line
  shows the parameter space for P$_{\rm out}$ allowed after
  considering all of the aforementioned criteria.  
\label{fig:fig2}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_M14MSun_outerorbit.pdf}
  \caption{Giant radius as a function of the mass of the core of the
    Roche-lobe filling outer star (dark blue curve).  Here we adopt
    a donor mass of 1.4\,\MSun, but for an 1.2\,\MSun\, donor the curve is
    quite similar (see dotted dark-blue curve).  The red squares in
    the curve show the parameters for which we performed more detailed
    gravitational-hydrodynamical simulations (see \S\,\ref{sims}).
    The horizontal dashed line shows the orbital separation of the
    observed twin BS 7782.  The initial triple in which it
    possibly formed must at least have been dynamically stable. The
    minimal orbital separation for the inner binary for which the
    triple is stable is given by the lower green coloured curve.
    Donors which are smaller than about 100\,\RSun\, (dark blue curve
    indicated with a$_{\rm stable}$) result in a dynamically unstable
    triple. The minimal core mass associated with a stable triple is
    then indicated by the left-most vertical dotted line.  The
    orbital separation at which the donor star overfills its Roche
    lobe is indicated with the light-blue curve. The top curve (brown)
    shows an estimate of the final orbital separation of the outer star,
    and therefore of the final orbit of the WD around the
    inner twin BSs.  For core masses $\apgt 0.5$\,\MSun\,
    the final orbital separation, after mass transfer, is smaller
    than the initial orbit.  Here we adopted an initial inner binary
    mass of (1.0+0.9)\,\MSun\, and a final twin BS mass of
    (1.4+1.4)\,\MSun.
\label{fig:tertiarymass_vs_size}}
\end{figure}

Adopting a mass for the tertiary star of $m_3 = 1.4$\,\MSun, we
can constrain the initial parameters for the inner binary as well as
the orbit of the outer star after mass transfer. We first calculate
the stellar radius as a function of core mass. In
figure\,\ref{fig:tertiarymass_vs_size} we present this relation
calculated using the {\tt SeBa} stellar evolution code
\cite{1996A&A...309..179P} (dark blue curve).  The interruption in
this curve, around a core mass of $m_{\rm core} \sim 0.5$\,\MSun\, is
a result of the evolution along the horizontal branch, where the core
of the star continues to grow but the radius actually shrinks.
Roche-lobe overflow in this phase is not expected to happen, because
it would already have happened in an earlier evolutionary state of the
donor star, when it was bigger.

Adopting masses for the inner binary $m_1=1.1$\,\MSun\, and
$m_2=0.9$\,\MSun\, we can calculate the outer orbital separation at
the onset of Roche-lobe overflow $a_{\rm out}$, and subsequently the
maximum orbital separation for the inner binary for which the orbit is
stable and a circumbinary disk can form. These two limits are
presented as the light blue and light green curves in
figure\,\ref{fig:tertiarymass_vs_size}.  The alotted region of
parameter space is then above the dashed horizontal line and to the
right of the vertical dotted line.

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after 
mass transfer.  The change in orbital separation due to
non-conservative mass transfer can be expressed in terms of the mass
of the outer star before and after mass transfer $m_3$ and $m'_3$,
respectively, the total mass in the inner binary before ($m_{\rm in}$)
and after accretion ($m'_{\rm in}$) and the amount of angular momentum
lost per unit mass $\eta \simeq 3$. Adopting the relation between the
orbital separation before mass transfer ($a$) and after mass transfer
($a'$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a' \over a} = \left( {m_3 m_{\rm in} \over m'_3 m'_{\rm in}} \right)^{-2}
  \left( {m_3 + m_{\rm in} \over m'_3 + m'_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top brown curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides a prediction for 
the current orbital separation of the WD around the twin
BS 7782. Tidal effects during mass tranfer
have probably circularized the orbit, although some slight eccentricity
due to turbulent motion in the outer layers of the donor star may have
induces a small $e \ll 0.1$ eccentricity.

Having limited parameter space for the formation of the twin BS 7782, we continue by performing a series of simulations to
investigate the accretion and changes to the inner orbits of triple
systems in this range of parameters.

\section{Numerical Simulations} \label{sims}

We perform simulations of a triple star system for which the outer
star overfills its Roche lobe while the inner binary remains
detached. The calculations start by evolving the three stars to the
same age, which is selected such that the outer-most star fills its
Roche lobe.  First order constraints for the initial conditions are
derived in the previous \S. In the following two sections we describe
how we set up these simulations and then discuss the results. The
calculations are performed using the Astrophysical Multipurpose
Software Environment using a combination of stellar evolution,
gravitational dynamics and hydrodynamics.

\subsection{Setting-up the simulations}

We adopt initial masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and
$m_{\rm 2} = 0.9$ M$_{\rm \odot}$ for the inner binary components, and
between $m_{\rm 3} = 1.2$ and $m_{\rm 3} = 1.4$ M$_{\rm \odot}$ for
the tertiary star.  We evolve the tertiary star using the MESA
stellar-evolution code \cite{2011ApJS..192....3P} to a radius of about
100\,\RSun\, and 150\,\RSun, at which point we assume it to overfill
it's Roche lobe (see red square in
figure~\ref{fig:tertiarymass_vs_size}).  We perform calculations for an
inner orbital separation of $a_{\rm in} = 0.10$\,au, $a_{\rm in} =
0.15$\,au and $a_{\rm in} = 0.20$\,au.  In total we performed 12
calculations at a resolution of 40k SPH particles and 12 at 80k.

The stellar-evolution model, including the structure, temperature and
composition profiles are turned into a smoothed-particles
representation using the module {\tt StellarModelInSPH} in AMUSE (see
chapter 4 in \cite{AMUSE}).  We follow the same procedure as described
in \cite{2014MNRAS.438.1909D} for simulating the future of the triple
system $\chi$ Tau (HD 97131) in which the outer-most star overfills
its Roche lobe and transfers mass to an inner binary.  After
generating the hydrodynamical representation of the donor star we
replace the stellar core by a point mass to prevent the majority of
the resolution to be confined in the star's central regions.  In a
following step we relax the star using the hydrodynamics solver. This
relaxation process is realized in 100 steps during which we reduce the
velocity dispersion of individual SPH particles to a glasses
structure \citep[see, for example, \S\,3.3 on page 40
  in][]{1994astro.ph.10043W}. During this procedure, the gaseous
envelope of the star tends to expand by about 20\%.  To determine the
radius of the evolving star we calculate Lagrangian radii and use the
distance to the stellar center which contains 90\% of its mass. From
this 90\% mass-radius relation we obtain the stellar radius and match
it with the Roche-lobe of the outer orbit.

With these parameters the orbital separation of the outer binary
becomes $\sim 250$\,\RSun for the 100\,\RSun donor star and about
430\,\RSun\, for the more evolved donor star.  We adopt the outer
orbit to be circular and in the plane of the inner binary.
%% We
%% performed several additional calculations for inclined orbits (choosing
%% a relative initial inclination $i = 9^\circ$, $i = 45^\circ$, $i =
%% 90^\circ$ and $i = 180^\circ$, but we wi
The low inclination is motivated by observed stellar triples composed
of low-mass stars to favour co-planar configurations
\citep{2018ApJ...854...44M}.
%% To test the effect of highly inclined
%% and retrogade orbits we include the $i = 90^\circ$ and $i = 180^\circ$
%% cases.
In figure~\ref{fig:topview_at_t0} we present a top view of the initial
conditions for one of these calculations.

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_BBSS_gas_M14Msun_A010au_t0002.pdf}
\caption{Top view of the simulated triple system in which the
  100\,\RSun\, outer star of 1.4\,\MSun\, over-fills its
  Roche-lobe.  The star is represented by 80000\,SPH particles and a
  core particle of $\sim 0.4$\,\MSun (black bullet). The two companion
  stars are represented as black bullets (to the right).  The inner
  binary is represented by the yellow and red bullets for, respectively, the
  1.1\,\MSun\, primary and 0.9\,\MSun\, secondary stars in a circular
  orbit of 0.1\,au. The 1.4\,\MSun\, giant star is presented to the
  right in a circular orbit of $\sim 250$\,\RSun\, in the plane of the
  inner binary.
\label{fig:topview_at_t0}}
\end{figure}

Roche-lobe overflow in triples is modelled using a coupled integrator
to follow the complex hydrodynamics of mass transfer from the
Roche-lobe filling outer star to the inner binary, while keeping track
of the gravitational dynamics of the stars.  The equations of motion
of the inner binary are solved using the symplectic direct N-body
integrator \texttt{Huayno} \citep{2012NewA...17..711P}. The
hydrodynamics are performed with the smoothed-particles hydrodynamics
code \texttt{Gadget2} \citep{2000ascl.soft03001S}, using an adiabatic
equation of state.  The two inner binary stars are treated as point
masses, but we allow them to accrete mass and angular momentum from
the gas liberated by the outer star.  This is realized using spherical
sink-particles that co-move with the mass points in the gravity
code. While the inner two stars accrete mass, they also accrete the
corresponding amount of angular momentum from the gas (see chapter 5
in \cite{AMUSE}).  The N-body integrator correctly accounts for this.
For the radius of the sink particles, we adopt $2 R_\odot$ for both
stars.

The N-body code, as well as the hydrodynamics solver, operate using
their own internal time-steps. The coupling between the two codes is
realized using the \texttt{Bridge} method in the AMUSE framework
\citep[see Sect.\.4.3.1 in][]{2013CoPhC.183..456P}.  This coupled
integrator is based on the splitting of the Hamiltonian, much in the
same way as is done with two different gravity solvers by
\cite{2007PASJ...59.1095F}. With the adopted scheme, the
hydrodynamical solver is affected by the gravitational potential of
its own particles, as well as the gravitational potential of the inner
binary. The hydrodynamics affects the orbits of the two inner stars. With
\texttt{Bridge} we realize a second order coupling between the
gravitational dynamics and the hydrodynamics.  The interval at which
the gravity and hydrodynamics interact via \texttt{Bridge} depends on
the parameters of the system we study, but typically we achieve
converged solutions when this time step is about a fraction of 1/100
of the inner binary orbital period.

\section{Results of the hydrodynamical simulations} \label{results}

to test the hypotheses that (1) the secondary in the inner binary
accretes more effectively than the primary star and to measure the
change to the inner orbit due to the Roche-lobe overflow of the outer
star, we perform a series of calculations in which we take the self
gravity and the hydrodynamical effects of the triple into account.
The results of these simulations are presented in
figure~\ref{fig:topview_at_t1000day} and
figure~\ref{fig:mass_vs_semimajor_axis}.  The first figure
(figure~\ref{fig:topview_at_t1000day}) shows the top view of the same
initial realization for which we presented the initial conditions in
figure~\ref{fig:topview_at_t0} but now at an age of 1091 days after
the onset of mass transfer. To guide the eye we add, to the left panel,
the equipotential surfaces in the orbital plane.

It is apparent that the mass transfer in the adopted triples leads to
a rather untidy evolution, in particular, because much of the donor
mass is lost through the second Lagrangian point to the right side of
the donor star in figure~\ref{fig:topview_at_t1000day}. Much mass is
also lost through the third Lagrangian point (to the left of the inner
binary). One remaining question is how much mass is eventually ejected
altogether from the triple system and is therefore not accreted to any
of the two inner stars. This value is hard to estimate from the
simulations, but an accretion efficiency of $\apgt 0.6$ is necessary to
make the scenario feasible. Over the time scale for which we performed
the calculations, this efficiency is reached, but it is not clear how
the system responds at later stages.

\begin{figure}[ht!]
  \includegraphics[width=0.49\columnwidth]{fig_BBSS_Lgas_M14Msun_A010au_t3360.pdf}
~  \includegraphics[width=0.51\columnwidth]{fig_BBSS_gas_M14Msun_A010au_t3360.pdf}
\caption{Same as figure~\ref{fig:topview_at_t0} but at an age of $t
  \simeq 1091\,days$ after the start of the simulation. Left panel
  shows the equipotential surfaces of the triple overplotted with the
  gas distribution, the right panel shows just the gas and the stars
  as bullets.
\label{fig:topview_at_t1000day}}
\end{figure}

The evolution of the inner orbit presented for several simulations in
figure~\ref{fig:mass_vs_semimajor_axis} is complicated.  This is
caused by the complex transport of mass, energy and angular momentum
through the accretion stream and throughout the system.  It is
therefore hard to quantify distinct trends in the evolution of the
triple system. In simulations of the response of an inner binary on
accretion from a circumbinary disk, \cite{2018arXiv181208175M}
conclude that the complexity of angular momentum transport between
the outer star and the accretion stream onto the individual inner stars, is
complicated and without clear trends. For most of our calculations we
agree with this statement, but in
figure~\,\ref{fig:mass_vs_semimajor_axis} we nevertheless present the
results of 6 of our calculations, three for a 1.2\,\MSun\, donor star and
three for a 1.4\,\MSun\, donor. The various coloured curves give
the resulting evolution of the inner orbit as a function of the total
mass in the inner binary. As the inner two stars accrete, the orbit shrinks 
for a 1.2\,\MSun\, donor. These systems
are expected to result in a contact binary, that eventually may merge
to form a single BS. The required evolution in order to
explain the observed twin BS 7782 is indicated by the
three black curves; the simulated path clearly deviates from
these. We, therefore, argue that a 1.2\,\MSun\, donor has difficulty
explaining the observed orbital separation of $\sim 0.13$\,au in BSS
7782.

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_orbital_evolution_for_12MSun.pdf}
  \includegraphics[width=\columnwidth]{fig_orbital_evolution_for_14MSun.pdf}
  \caption{Evolution of the total mass of the inner binary as a
    function of the orbital separation for four calculations with
    somewhat different initial conditions (see the legend).  The initial
    binary shown by the orange curve is presented in
    fig.\,\ref{fig:topview_at_t0}, and in
    fig.\,\ref{fig:topview_at_t1000day} we present the final
    conditions.
\label{fig:mass_vs_semimajor_axis}}
\end{figure}

In the lower panel in figure~\ref{fig:mass_vs_semimajor_axis} we
present the evolution of the orbit for the 1.4\,\MSun\, donor. A more
massive donor is more likely to be able to produce a twin BS system, simply because there is more mass available in the stellar
envelope. On the other hand, a more massive donor may also result in a
lower accretion efficiency, because the thermal time scale of the
stellar envelope is smaller than for a lower mass donor at the same
stellar radius. A shorter thermal time-scale eventually leads to
a higher mass transfer rate, which again may lead to a lower accretion
efficiency. However, the larger mass budget in the envelope appears to
compensate.

The orbit of the inner binary expands in these cases as a result of
accretion onto the inner two stars. All three cases
presented in figure~\ref{fig:mass_vs_semimajor_axis} expand at about
the same rate. Consequently, the inner binaries that start with $a =
0.15$\,au and $a=0.20$\,au eventually become dynamically unstable.
The binary with an initial separation of $0.10$\,au expands to reach a
separation of about 0.126--0.145\,au for final masses for the inner two stars
of 1.4\,\MSun, which is consistent with the observed twin BS 7782. In our simulations the eccentricity of the inner
binary grows to about $e \simeq 0.0028$.

With the accretion of mass, both stars in the inner binary also
accrete angular momentum.  By the end of the simulation the spins of
the two BSs are aligned along the orbital angular momentum
axis with an angle of $90.0^\circ$ for the primary star and
$93.4^\circ$ for the secondary star with respect to the argument of
pericenter of the inner orbit.  By the end of the simulations the spin
of the primary is about 50.5 rotations per day, and 41.5 rotations per
day for the secondary star. But by this time both stars have accreted
only a small portion of mass and we expect that the final spins of
both stars will be close to break-up (ignoring any spin-down from
magnetic braking effects).

\section{Discussion} \label{sect:discussion}

In this paper, we consider the formation of twin BSs in tight binaries.
These systems may form through mass transfer from an outer Roche-lobe
filling tertiary star. Once this star ascends the giant branch, part
of its envelope is transferred to the inner binary, and accreted by
the two inner stars which are still on the MS.

As illustrated via SPH simulations, the mass transfer stream forms a
circumbinary disk, from which the inner binary stars accrete, driving
the inner binary toward a mass ratio close to unity.  Our simulations
suggest that the inner binary orbital separation can decrease or
expand depending on the details of the transfer of mass and angular
momentum.  More work is certainly needed in order to fully understand 
mass transfer in triples.
%Regretfully, we do not really understand the details of this
%process, but then this process is hardly understood in binary
%evolution also.
%NL: I edited the above to be a kinder to us.

We summarize the results of these simulations as follows: for a
1.2\,\MSun\, tertiary donor mass, we expect the inner two stars to
eventually merge and form a single BS. This reduces the
system to a binary with a primary BS and an outer WD in a relatively wide orbit. Such a BS will
distinguish itself from other BSs by potentially
being more than twice the turn-off mass in a star cluster.  An
example could be the $2.9\pm0.2$\,\MSun\, BS S1237 in the
Galactic cluster M67 \cite{2016ApJ...832L..13L}. It is the primary of
a $\sim 698$\,day binary with an eccentric orbit of $\sim 0.10$.

With an original outer star of mass $\sim 1.4$\,\MSun, the inner orbit
tends to expand. This eventually leads
to a dynamically unstable system resulting either in a collision or in
the ejection of (probably) the lowest mass star. This evolution could
result in a single ejected blue straggler, with the other BS left in
a relatively close and eccentric orbit with a WD (the
left-over core of the tertiary star).  These ``imposter'' BS-WD binaries 
would in principle mimic what is expected theoretically for BSs formed 
from mass transfer in binary stars.

When we adopt an inner orbit of $0.10$\,au\, the expansion eventually
matches the observed orbital separation (i.e., $0.13$\,au) of the observed
twin BS 7782 and the observed masses of the two stars.

In order to study the T-tauri binaries V4046 Sgr and DQ Tau,
\cite{2011MNRAS.413.2679D} perform a series of 2D hydrodynamical
simulations of circumbinary disks.  These authors studied the two
observed T-tauri systems V4046 Sgr and DQ~Tau, to which we compare our
results here.  For V4046 Sgr, for which the two stars have comparable
masses as in our calculation for a circular orbit with a period of
only 2.4 days, they find that the inner binary accretes at a rate of
$\sim 0.028$\,\MSun/Myr.  For DQ~Tau, which is composed of lower-mass
stars ($m_1 = m_2 \simeq 0.55$\,\MSun) in an eccentric ($e\simeq
0.556$) orbit of $\sim 15.8$\,days, they find an accretion rate onto
the inner binary of $\sim 0.027$\,\MSun/Myr.  These values are in the
same range as in our calculations, which results in an accretion rate
for the inner binary of 0.027--0.058\,\MSun/Myr (i.e., the average
measured over a period of about 3000 days in our simulations).
Interestingly, however, \cite{2011MNRAS.413.2679D} find that the
primary star in V4046 Sgr accretes at an 8\% higher rate than the
secondary star, whereas in our case the secondary star accretes at a
higher rate than the primary star by about 1\% to 12\%.  Higher
accretion rates in the secondary star are realized for eccentric and
retrogade inner orbits.

%NL2: Huh.  This is interesting.  Is the physical reason for it that
%apocenter for the secondary tends to be closer to the circumbinary
%disk?  I don't get why retrograde wins though, since naively I'd have
%thought prograde should lead to lower relative velocities between the
%secondary and the gas flow, and hence to higher accretion rates.
%What am I missing?  Is there momentum cancellation via colliding gas,
%shocks, etc. I am missing somewhere?
%% SPZ: I think that we do not have to add these extra simulations. I
%% found the result sufficiently confusing to check and check
%% again. Maybe we write another paper on these results, or a
%% follow-up on accretion in triples. But maybe it is best to stick to
%% the observed facts, and not try to explore parameter space in
%% extensio. What do you think, Nathan?  At the moment, this last
%% paragraph is a bit of an after though, but maybe it is worth
%% mentioning that our results are consistent with those of others...

\section{Summary} \label{sect:conclusions}

In this paper, we propose a formation scenario for twin equal-mass blue stragglers 
in tight binaries, as observed for Binary 7782 in the old OC
NGC 188.  The proposed scenario involves mass transfer from an evolved
outer tertiary companion, which is accreted by the inner binary via a
circumbinary disk.  Our scenario makes several predictions for the
observed properties of a hypothetical outer triple companion, now a
WD.  These are:

\begin{enumerate}

\item For the predicted outer tertiary orbit, the initial 
  orbital period should lie between 220 days $\lesssim$ P$_{\rm out}$
  $\lesssim$ 1100 days, assuming initial masses for the inner binary
  components of $m_{\rm 1} = 1.1$ M$_{\odot}$ and $m_{\rm 2} = 0.9$
  M$_{\odot}$ and an initial outer tertiary mass of $m_{\rm 3} = 1.4
  $M$_{\odot}$.  Hence, the present-day orbital period should be close to this range.

\item Larger final WD masses, and hence larger core masses for the donor at
  the time of mass transfer should correspond to larger final outer
  orbital periods for the tertiary.  This is because the Roche radius
  is larger for larger outer orbital periods, such that the donor must
  evolve to larger radii, and hence core masses, before the onset of
  mass transfer.

\item For the inner binary, the rotational axes of both the BSs should
  be aligned with each other and the orbital plane of the outer
  tertiary WD.  This is because accretion onto the BS progenitors
  proceeds via an accretion disk, that forms at the circularization
  radius and that has an orbital plane aligned with that of the outer
  tertiary.

\item The BSs in the inner binary should have roughly equal masses,
  independent of their initial masses.  This is because it is the
  lowest mass object that typically accretes the fastest, since its
  orbital velocity and distance relative to the circumbinary disk is
  typically the lowest
  \citep[e.g.][]{1995MNRAS.277.1491K,1997MNRAS.285...33B,2000MNRAS.314...33B}.
  This quickly brings the mass ratio toward unity.  This predicts that
  the initially lower mass MS star should accrete the most, and should
  hence be enriched more by any accreted material.  This could be
  observable in the surface layers of a radiative star.  If the donor
  is an RGB star, the accretor will be enriched in mostly carbon,
  oxygen and helium.  If the donor is an AGB star, it will be enriched
  in mostly s-process elements.

\item Twin BSs in compact binaries formed from the mechanism proposed here 
  should be more frequent in younger clusters with ages $\lesssim$ 4-6
  Gyr.  This is because clusters with a MS turn-off mass
  $\lesssim$ 1.2 M$_{\odot}$ have convective envelopes
  \citep[e.g.][]{1991ApJS...76...55I,2009pfer.book.....M}, and a radiative envelope for the
  donor in a mass transferring binary ensures stable accretion on to
  the accretor.  \simon{Can we quantify this?}

\end{enumerate}

We emphasize in closing that our choice for the initial mass of the
outer tertiary is critical since it ensures that the donor star during
the mass transfer process will have a radiative envelope
\citep[e.g.][]{2009pfer.book.....M}.  In turn, this ensures that the
mass transfer will be maximally conservative, such that the accretion
stream will be maximally stable, accreting at a stable and roughly
constant rate \citep[e.g.][]{1991ApJS...76...55I}.

\acknowledgments

N.W.C.L. acknowledges support from a Kalbfleisch Fellowship at the
American Museum of Natural History.  SPZ would like to thank Norm
Murray and CITA for their hospitality during my long-term visit.  This
work was supported by the Netherlands Research School for Astronomy
(NOVA), NWO (grant \# 621.016.701 [LGM-II]).

\bibliographystyle{apj}
\bibliography{BSS7782}

%\input /home/spz/Latex/lib/bib/references
%\bibliographystyle{/home/spz/Latex/lib/styles/elsevier/elsarticle-num} 
%\bibliography{references}

\end{document}